---
title: "SCS 3547 - Intelligent Agents & Reinforcement Learning"
author: "Bill Zizek"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

# Assignment 2 

## Objective

Take a copy of the WumpusWorld solution found [here](https://github.com/rlegendi/wumpus-prolog/blob/master/wumpus.pl). This code can be run [here](https://swish.swi-prolog.org/).

Write an explanation of how each **rule** in the `Prolog` program works. Rule headers in `Prolog` can be identified by looking for `:-`. For example:

```{r prolog example}
father(X,Y) :- parent(X,Y), male(X).
```

X is father of Y **if** X is a parent of Y **and** male.

Further syntax explanations available online in Prolog books/tutorials. 

## Explanations
### *Code chunks will be displayed, followed by the associated explanation*

```{r declaring dynamic methods}
:- dynamic ([
	     agent_location/1,
	     gold_location/1,
	     pit_location/1,
	     time_taken/1,
	     score/1,
	     visited/1,
	     visited_cells/1,
	     world_size/1,
	     wumpus_location/1,
             isPit/2,
             isWumpus/2,
             isGold/2
	    ]).
```

Here, we are declaring dynamic methods for various facets of WumpusWorld including locations, scores, world size, and more. Prolog has both static and dynamic procedures. An autonomous agent's strategy has been coded for in the dynamic WumpusWorld - so as such a static procedure which would never change would not be suitable. We are simply declaring identifiers within a dynamic method at this stage, so these identifiers will be defined below. The numbers included indicate how many arguments the identifier will take. 

Let's get started!

```{r start game}
start :-
    format('Initializing started...~n', []),
    init,
    format('Let the game begin!~n', []),
    take_steps([[1,1]]).
```

`start` the game (no arguments) **if** the predicate `format` printing some text is successful with an empty list **and** the `init` functor can be called to set up the WumpusWorld (more on this below) with no arguments necessary **and** the predicate `format` printing some text is successful with an empty list **and** a fact called `take_steps` to modulate location (more on this below) is called using a list of lists to set us in the starting position 1,1.

```{r scheduling simulation}
step_pre(VisitedList) :-
    agent_location(AL),
    gold_location(GL),
    wumpus_location(WL),
    score(S),
    time_taken(T),

    ( AL=GL -> writeln('WON!'), format('Score: ~p,~n Time: ~p', [S,T])
    ; AL=WL -> format('Lost: Wumpus eats you!~n', []),
               format('Score: ~p,~n Time: ~p', [S,T])
    ; take_steps(VisitedList)
    ).

take_steps(VisitedList) :-
    make_percept_sentence(Perception),
    agent_location(AL),
    format('I\'m in ~p, seeing: ~p~n', [AL,Perception]),

    update_KB(Perception),
    ask_KB(VisitedList, Action),
    format('I\'m going to: ~p~n', [Action]),

    update_time,
    update_score,

    agent_location(Aloc),
    VL = [Aloc|VisitedList],
    standing,
    step_pre(VL).
```

First we define the rule `step_pre` using `VisitedList` wherein we check if the agent, gold, and wumpus location warrant an update of the game status (win/loss), to stop the game (elsewhere) and print the outcome to the console if no next move is required based on where the agent has been. The rule is that if the agent, gold, and wumpus locations (+ arguments) equal a set of defined conditions (such as agent location matching the gold location or wumpus location) - the predicate `format` will print some text contingent on the outcome alongside the matching score and time taken. If the agent is not at the gold or wumpus, the `;` is a disjunction operator in prologue so the agent will take more steps. This historical list of locations is what declares the end of the game. 

Now we define our `take_steps` rule again using the `VisitedList` argument. First the environment perception of neighboring nodes is made (we define this later - breeze, stench, and glitter). In addition to this perception, we bind this with the agent location and print it to the console so we know where the current location is and what's around. This perception is what an agent uses to make a decision of where to move next. We add this perception information to our knowledgebase tracking information about the wumpus, a pit, and if the move is permitted. Then we print where we're going to the console, update the time + score, update our visited list, update the standings, and take the step based on the `step_pre` rule we just defined!

```{r updating states}
update_time :-
    time_taken(T),
    NewTime is T+1,
    retractall( time_taken(_) ),
    assert( time_taken(NewTime) ).

update_score :-
    agent_location(AL),
    gold_location(GL),
    wumpus_location(WL),
    update_score(AL, GL, WL).

update_score(P) :-
    score(S),
    NewScore is S+P,
    retractall( score(_) ),
    assert( score(NewScore) ).

update_score(AL, AL, _) :-
    update_score(1000).

update_score(_,_,_) :-
    update_score(-1).

update_agent_location(NewAL) :-
    retractall( agent_location(_) ),
    assert( agent_location(NewAL) ).

is_pit(no,  X) :-
    \+ pit_location(X).
is_pit(yes, X) :-
    pit_location(X).
```

```{r display standings}
standing :-
    wumpus_location(WL),
    gold_location(GL),
    agent_location(AL),

    ( is_pit(yes, AL) -> format('Agent was fallen into a pit!~n', []),
      fail
    ; stnd(AL, GL, WL)
      %\+ pit_location(yes, Al),
    ).

stnd(_, _, _) :-
    format('There\'s still something to do...~n', []).

stnd(AL, _, AL) :-
    format('YIKES! You\'re eaten by the wumpus!', []),
    fail.

stnd(AL, AL, _) :-
    format('AGENT FOUND THE GOLD!!', []),
    true.
```

```{r perception}
make_perception([_Stench,_Bleeze,_Glitter]) :-
    agent_location(AL),
    isStinky(AL),
    isBleezie(AL),
    isGlittering(AL).

test_perception :-
	make_percept_sentence(Percept),
	format('I feel ~p, ',[Percept]).

make_percept_sentence([Stench,Bleeze,Glitter]) :-
	smelly(Stench),
	bleezy(Bleeze),
	glittering(Glitter).
```

```{r initializing}
init :-
    init_game,
    init_land_fig72,
    init_agent,
    init_wumpus.

init_game :-
    retractall( time_taken(_) ),
    assert( time_taken(0) ),

    retractall( score(_) ),
    assert( score(0) ),

    retractall( visited(_) ),
    assert( visited(1) ),

    retractall( isWumpus(_,_) ),
    retractall( isGold(_,_) ),

    retractall( visited_cells(_) ),
    assert( visited_cells([]) ).

init_land_fig72 :-
    retractall( world_size(_) ),
    assert( world_size(4) ),

    retractall( gold_location(_) ),
    assert( gold_location([3,2]) ),

    retractall( pit_location(_) ),
    assert( pit_location([4,4]) ),
    assert( pit_location([3,3]) ),
    assert( pit_location([1,3]) ).

init_agent :-
    retractall( agent_location(_) ),
    assert( agent_location([1,1]) ),

    visit([1,1]).

init_wumpus :-
    retractall( wumpus_location(_) ),
    assert( wumpus_location([4,1]) ).

visit(Xs) :-
    visited_cells(Ys),
    retractall( visited_cells(_) ),
    assert( visited_cells([Ys|Xs]) ).
```

```{r Perceptors}
%%% Institiation error!!!

%adj(X,Y) :-
%    world_size(WS),
%    ( X is Y+1, Y   < WS
%    ; X is Y-1, Y-1 > 0
%    ).

adj(1,2).
adj(2,1).
adj(2,3).
adj(3,2).
adj(3,4).
adj(4,3).

adjacent( [X1, Y1], [X2, Y2] ) :-
    ( X1 = X2, adj( Y1, Y2 )
    ; Y1 = Y2, adj( X1, X2 )
    ).

%adjacent([X1,Y],[X2,Y]) :-
%    adj(X1,X2).

%adjacent([X,Y1],[X,Y2]) :-
%    adj(Y1,Y2).

isSmelly(Ls1) :-
    wumpus_location( Ls2 ),
    adjacent( Ls1, Ls2 ).

isBleezy(Ls1) :-
    pit_location( Ls2 ),
    adjacent( Ls1, Ls2 ).

isGlittering( [X1, Y1] ) :-
    gold_location( [X2, Y2] ),
    X1 = X2,
    Y1 = Y2.

bleezy(yes) :-
    agent_location(AL),
    isBleezy(AL).
bleezy(no).

smelly(yes) :-
    agent_location(AL),
    isSmelly(AL).
smelly(no).

glittering(yes) :-
    agent_location(AL),
    isGlittering(AL).
glittering(no).
```

```{r knowledge base}
update_KB( [Stench,Bleeze,Glitter] ) :-
    add_wumpus_KB(Stench),
    add_pit_KB(Bleeze),
    add_gold_KB(Glitter).

% if it would be 'yes' -> it would mean the player is eaten ;]
add_wumpus_KB(no) :-
    %agent_location(L1),
    %adjacent(L1, L2),
    %assume_wumpus(no, L2).
    agent_location([X,Y]),
    world_size(_),

    % Checking needed!!
    % adj will freeze for (4,_) !!

    Z1 is Y+1, assume_wumpus(no,[X,Z1]),
    Z2 is Y-1, assume_wumpus(no,[X,Z2]),
    Z3 is X+1, assume_wumpus(no,[Z3,Y]),
    Z4 is X-1, assume_wumpus(no,[Z4,Y]).

add_pit_KB(no) :-
    agent_location([X,Y]),
    Z1 is Y+1, assume_pit(no,[X,Z1]),
    Z2 is Y-1, assume_pit(no,[X,Z2]),
    Z3 is X+1, assume_pit(no,[Z3,Y]),
    Z4 is X-1, assume_pit(no,[Z4,Y]).

% Checking needed!! If its not already in the KB !!!
add_pit_KB(yes) :-
    agent_location([X,Y]),
    Z1 is Y+1, assume_pit(yes,[X,Z1]),
    Z2 is Y-1, assume_pit(yes,[X,Z2]),
    Z3 is X+1, assume_pit(yes,[Z3,Y]),
    Z4 is X-1, assume_pit(yes,[Z4,Y]).

add_gold_KB(no) :-
    gold_location(GL),
    assume_gold(no, GL).

add_gold_KB(yes) :-
    gold_location([X1,Y1]),
    agent_location([X2,Y2]),
    X1 = X2, Y1 = Y2,
    assume_gold(yes, [X1,Y1]).

assume_wumpus(no, L) :-
    retractall( isWumpus(_, L) ),
    assert( isWumpus(no, L) ),
    format('KB learn ~p - no Wumpus there!~n', [L]).

assume_wumpus(yes, L) :-
    %wumpus_healthy, % Will be included ...
    retractall( isWumpus(_, L) ),
    assert( isWumpus(yes, L) ),
    format('KB learn ~p - possibly the Wumpus is there!~n', [L]).

assume_pit(no, L) :-
    retractall( isPit(_, L) ),
    assert( isPit(no, L) ),
    format('KB learn ~p - there\'s no Pit there!~n', [L]).

assume_pit(yes, L) :-
    retractall( isPit(_, L) ),
    assert( isPit(yes, L) ),
    format('KB learn ~p - its a Pit!~n', [L]).

assume_gold(no, L) :-
    retractall( isGold(_, L) ),
    assert( isGold(no, L) ),
    format('KB learn ~p - there\'s no gold here!~n', [L]).

assume_gold(yes, L) :-
    retractall( isGold(_, L) ),
    assert( isGold(yes, L) ),
    format('KB learn ~p - GOT THE GOLD!!!~n', [L]).

permitted([X,Y]) :-
    world_size(WS),
    0 < X, X < WS+1,
    0 < Y, Y < WS+1.

ask_KB(VisitedList, Action) :-
    isWumpus(no, L),
    isPit(no, L),
    permitted(L),
    not_member(L, VisitedList),
    update_agent_location(L),
    Action = L.
```

```{r utils}
not_member(_, []).
not_member([X,Y], [[U,V]|Ys]) :-
    ( X=U,Y=V -> fail
    ; not_member([X,Y], Ys)
    ).
```

